---
title: "P53 library - activity computation - all data combined"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute - van Steensel lab"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output: 
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

---


### Introduction
~18,000 barcoded P53 reporters were probed in MCF7 P53WT/KO cells and stimulated with Nutlin-3a. I previously processed the raw sequencing data, quantified the pDNA data and normalized the cDNA data. In this script, a detailed dissection of the reporter activities will be carried out to understand how P53 drives transcription and to identify the most sensitive P53 reporters.


---

## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```

### Libraries 

```{r setup, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(maditr)
library(tibble)
library(pheatmap)
library(ggpubr)
library(ggbeeswarm)
library(ggforce)
library(viridis)
library(plyr)
library(cowplot)
library(gridExtra)
library(GGally)
library(readr)
library(stringr)
library(tidyr)
library(ROCR)
library(plotly)
library(randomForest)
library(glmnet)
library(glmnetUtils)
library(jtools)
```

---

### Functions

```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}


cbind.fill <- function(...){
    nm <- list(...) 
    nm <- lapply(nm, as.matrix)
    n <- max(sapply(nm, nrow)) 
    do.call(cbind, lapply(nm, function (x) 
        rbind(x, matrix(, n-nrow(x), ncol(x))))) 
}

# Function to substring the right part of the motif
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


# Function to load PWM matrix
get_pwm_feature_matrix <- function(motif_meta_fn, fimo_fn, db = 2) {

  # validate args
  valid_dbs <- 1:2
  if(!db %in% valid_dbs)
    stop('Invalid db (database version). Please use db=1 (maintained for backward compatibility only) or db=2')

  # db=1 is maintained for backward compatibility only
  if(db == 1) {

    # read in motif metadata
    motif_meta    <- read.csv(motif_meta_fn)

    # check whether motif metadata contain essential annotations
    if(!all(c('PWM.ID', 'Cognate.TF') %in% colnames(motif_meta))) {
      message('The motif metadata file does not contain the essential columns PWM.ID and Cognate.TF')
    }

    motif_minimal <- motif_meta[, c('PWM.ID', 'Cognate.TF')]

    # load fimo output --> extract motif id, sequence id and p-value
    df <- read.table(fimo_fn)
    df <- df[, c(1, 2, 7)]

    colnames(df) <- c('PWM.ID', 'seqid', 'pval')

    # add TF id
    df <- merge(df, motif_minimal, by = 'PWM.ID')

    # group motif hits by sequence id
    l <- split(df, df[['seqid']])

    # multiple PWM and multiple hits possible. Reduce hits to one per TF, keeping best p-val only
    l <- lapply(l, function(x) {
      x_by_tf <- split(x, x[['Cognate.TF']], drop = TRUE)
      x_by_tf <- lapply(x_by_tf, function(y) y[which.min(y$pval), ])
      do.call('rbind', x_by_tf)
    })

    # initialize feature matrix
    n_tf          <- motif_minimal[['Cognate.TF']] %>%
      unique %>%
      length
    n_seq         <- length(l)
    pwm           <- matrix(1, nrow = n_seq, ncol = n_tf)
    colnames(pwm) <- (motif_minimal[['Cognate.TF']] %>% unique)

    # replace :: from names of composite motifs
    colnames(pwm) <- str_replace_all(colnames(pwm), '::', '_')

    # fill in feature matrix
    for(i in 1 : n_seq) {
      pwm[i, l[[i]][['Cognate.TF']]] <- l[[i]]$pval
    }

    # -log10 transform
    pwm           <- -1 * log10(pwm)

    # coerce to tib and return
    tib_fimo <- as_data_frame(pwm) %>%
      mutate(id = names(l))
      dplyr::select(id, everything())

  }

  # db = 2 (default)
  else {

    # load metadata
    tib_meta    <- read_csv(motif_meta_fn) %>%
      # extract tf symbol from motif id (Cognate_TF unsafe, it can be empty) and replace :: occurrences
      mutate(tf_symbol = str_remove(ID, '_[0-9]*'),
             tf_symbol = str_replace(tf_symbol, '::', '_')) %>%
      dplyr::select(motif_id = `PWM ID`, tf_symbol)

    # load fimo results
    tib_fimo <- read_tsv(fimo_fn) %>%
      # extract motif id, sequence id and p-value
      dplyr::select(motif_id, sequence_name, pval = `p-value`)

    # add tf symbol to fimo results
    tib_fimo <- tib_fimo %>%
      left_join(tib_meta, by = 'motif_id') %>%
      # remove hits with missing motif id (composite pwms)
      filter(!is.na(tf_symbol))

    # select best hit for each motif and sequence
    tib_fimo <- tib_fimo %>%
      group_by(sequence_name, tf_symbol) %>%
      dplyr::slice(which.min(pval)) %>%
      ungroup()

    # spread into feature matrix
    tib_fimo <- tib_fimo %>%
      mutate(pval = -1 * log10(pval)) %>%
      dplyr::select(-motif_id) %>%
      spread(key = tf_symbol, value = pval, fill = 0, drop = TRUE) %>%
      # perform cosmetics on the id
      mutate(id = sequence_name) %>%
      dplyr::select(-c(sequence_name)) %>%
      dplyr::select(id, everything())

  }

  return(tib_fimo)

}

# From Fede:
# ggpairs custom functions
corColor <- function(data, mapping, color = I("black"), sizeRange = c(1, 3), ...) {

  x   <- eval_data_col(data, mapping$x)
  y   <- eval_data_col(data, mapping$y)
  r   <- cor(x, y, "pairwise.complete.obs")
  rt  <- format(r, digits = 3)
  tt  <- as.character(rt)
  cex <- max(sizeRange)

  # helper function to calculate a useable size
  percent_of_range <- function(percent, range) {
    percent * diff(range) + min(range, na.rm = TRUE)
  }

  # plot correlation coefficient
  p <- ggally_text(label = tt, mapping = aes(), xP = 0.5, yP = 0.5,
                   size = I(percent_of_range(cex * abs(r), sizeRange)), color = color, ...) +
    theme(panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())

  corColors <- RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")[2:6]

  if (r <= boundaries[1]) {
    corCol <- corColors[1]
  } else if (r <= boundaries[2]) {
    corCol <- corColors[2]
  } else if (r < boundaries[3]) {
    corCol <- corColors[3]
  } else if (r < boundaries[4]) {
    corCol <- corColors[4]
  } else {
    corCol <- corColors[5]
  }

  p <- p +
    theme(panel.background = element_rect(fill = corCol))

  return(p)
}


# Custom ggplot2 themes
theme_classic_lines <- function() {
  theme_pubr(border = T, base_family = "Arial", legend = "top") +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.1),
                  strip.background = element_rect(fill = "#ced4da"))
    
}

theme_classic_lines_45 <- function() {
  theme_pubr(border = T, base_family = "Arial", legend = "top", x.text.angle = 45) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.1),
                  strip.background = element_rect(fill = "#ced4da"))
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T, base_family = "Arial", legend = "top", x.text.angle = 90) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.1),
                  strip.background = element_rect(fill = "#ced4da"))
}

theme_set(theme_classic_lines())

ggplot_cust <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)

## save favorite colors
colors_diverse <- c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51")
colors_diverse_2 <- c("#CB997E", "#DDBEA9", "#FFE8D6", "#B7B7A4", "#A5A58D", "#6B705C")
colors_continous <- c("#CAD2C5", "#84A98C", "#52796F", "#354F52", "#2F3E46")
colors_continous_2 <- c("#F8F9FA", "#E9ECEF", "#DEE2E6", "#CED4DA", "#ADB5BD", "#6C757D", "#495057", "#343A40", "#212529")
reds <- c("#DD6B48", "#E38569", "#E7A08A", "#EDBBAB", "#EFC9BC")
```

---

### Load data

```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# Import processed bc counts from the preprocessing step
cDNA_df <- read.csv("/DATA/usr/m.trauernicht/projects/P53_reporter_scan/data/gcf6502/results/mt20210819_reporter_activity_filt.csv", header = T) 
cDNA_df <- cDNA_df[!is.na(cDNA_df$tf),]
cDNA_df$activity[is.na(cDNA_df$activity)] <- 0

# Remove outliers and recalculate mean activities
cDNA_df <- cDNA_df %>%
  filter(outlier == "No") %>%
  mutate(reporter_activity = ave(reporter_activity, reporter_id, sample, FUN = function(x) mean(x)),
         log_reporter_activity = log2(reporter_activity)) %>%
  dplyr::select(-reporter_activity_sd)


# Change names
cDNA_df$tf <- as.character(cDNA_df$tf)
cDNA_df$condition[cDNA_df$condition == "MCF7_PFTa"] <- "MCF7_WT_PFTa"
cDNA_df$condition[cDNA_df$condition == "MCF7_DMSO"] <- "MCF7_WT_DMSO"
cDNA_df$condition[cDNA_df$condition == "MCF7_Nutlin"] <- "MCF7_WT_Nutlin"

# Remove gcf6501 KO data because it looks like P53 regained activity there
cDNA_df <- cDNA_df %>%
  filter(sample != "MCF7_KO_DMSO_r1_gcf6501", sample != "MCF7_KO_DMSO_r2_gcf6501")
cDNA_df <- cDNA_df %>%
  filter(sample != "MCF7_KO_Nutlin_r1_gcf6501", sample != "MCF7_KO_Nutlin_r2_gcf6501")
cDNA_df$sample[cDNA_df$sample == "MCF7_KO_DMSO_r1_gcf6412"] <- "MCF7_KO_DMSO_r1_gcf6501"
cDNA_df$gcf[cDNA_df$sample == "MCF7_KO_DMSO_r1_gcf6501"]  <- "gcf6501"

# Check if the conditions are scaled correctly
ggplot_cust(cDNA_df %>% 
         filter(neg_ctrls == "Yes", str_detect(tf, "53"), str_detect(condition, "MCF")) %>% 
         dplyr::select(tf, condition, reporter_activity, reporter_id, gcf, promoter) %>% 
         unique(), 
       aes(x = condition, y = reporter_activity, color = tf)) + 
  geom_quasirandom(dodge.width = 0.75) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  facet_grid(promoter~gcf)



### In the barcode-preprocessing script I already scaled the data so that the negative controls would have the same counts in all conditions
# Now I want to compute the enrichment over the inactive ones to center the data around 1 (1 = inactive)
cDNA_df_neg <- cDNA_df %>%
  dplyr::select(tf, condition, reporter_activity, reporter_id, motif_id, gcf, promoter) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(motif_id == "4_4_4_4", promoter == "random") %>%
  unique() %>%
  group_by(condition, gcf) %>%
  mutate(reporter_activity = mean(reporter_activity)) %>%
  ungroup() %>%
  dplyr::select(-reporter_id) %>%
  unique() %>%
  dplyr::select("background_activity" = reporter_activity, condition, gcf) %>%
  unique()

cDNA_df <- cDNA_df %>%
  mutate(tf = gsub("_.*", "", tf))
cDNA_df <- merge(cDNA_df, cDNA_df_neg)
cDNA_df <- cDNA_df %>%
  mutate(reporter_activity_norm = reporter_activity / background_activity,
         activity_norm = activity / background_activity)

# Assign which pDNA library was used for which gcf
cDNA_df$pDNA_library <- "lib_1"
#cDNA_df$pDNA_library[cDNA_df$gcf == "gcf6412" | cDNA_df$gcf == "gcf6501"] <- "lib_2"
cDNA_df$pDNA_library[cDNA_df$gcf == "gcf6501"] <- "lib_2"

# # Remove outlier sample: MCF7_WT_DMSO_r1_gcf6412
# cDNA_df <- cDNA_df %>%
#   filter(sample != "MCF7_WT_DMSO_r1_gcf6412")

# Compute means per sample
cDNA_df$reporter_activity_condition <- ave(cDNA_df$reporter_activity_norm, cDNA_df$condition, cDNA_df$reporter_id, cDNA_df$pDNA_library,
                                           FUN = function(x) mean(x))
cDNA_df$reporter_activity_sample <- ave(cDNA_df$reporter_activity_norm, cDNA_df$sample, cDNA_df$reporter_id, 
                                           FUN = function(x) mean(x))

# Check if the conditions are scaled correctly
ggplot_cust(cDNA_df %>% 
         filter(tf == "P53", neg_ctrls == "Yes", str_detect(condition, "MCF")) %>% 
         dplyr::select(tf, sample, reporter_activity_sample, reporter_id, gcf, promoter) %>% 
         unique(), 
       aes(x = sample, y = reporter_activity_sample, color = tf)) + 
  geom_quasirandom(dodge.width = 0.75) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  facet_grid(gcf~promoter)
```

---

## Figure 1: Characterize P53 activities per condition

Aim: I want to characterize the reporter activity distributions in the tested conditions. Does Nutlin boost P53 reporter activity and is P53 inactive in the KO cells?
```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Plot activities in the different conditions
cDNA_df_cor_mcf <- cDNA_df %>%
  filter(tf == "P53", motif_id != "4_4_4_4", promoter != "random", str_detect(reporter_id, "random", negate = T)) %>%
  filter(str_detect(condition, "MCF7_WT"), pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T)) %>%
  dplyr::select(reporter_id, reporter_activity_sample, condition, reporter_activity_condition) %>%
  unique()

for (i in unique(cDNA_df_cor_mcf$reporter_id)) {
  if (nrow(cDNA_df_cor_mcf[cDNA_df_cor_mcf$reporter_id == i,]) == 4) {
  cDNA_df_cor_mcf$pval[cDNA_df_cor_mcf$reporter_id == i] <- 
    t.test(cDNA_df_cor_mcf$reporter_activity_sample[cDNA_df_cor_mcf$condition == "MCF7_WT_DMSO" & cDNA_df_cor_mcf$reporter_id == i],
           cDNA_df_cor_mcf$reporter_activity_sample[cDNA_df_cor_mcf$condition == "MCF7_WT_Nutlin" & cDNA_df_cor_mcf$reporter_id == i])$p.value
  }
  else {
    cDNA_df_cor_mcf$pval[cDNA_df_cor_mcf$reporter_id == i] <- NA
  }
}

cDNA_df_cor_mcf <- cDNA_df_cor_mcf %>%
  dplyr::select(-reporter_activity_sample) %>%
  unique() %>%
  spread(condition, reporter_activity_condition) %>%
  mutate(log2dif = log2(MCF7_WT_Nutlin / MCF7_WT_DMSO),
         log10pval = -log10(pval))

ggplot_cust(cDNA_df_cor_mcf,
            aes(x = log2dif, y = log10pval)) +
  geom_point()


cDNA_df_cor_mcf$sensitive <- "No"
cDNA_df_cor_mcf$sensitive[cDNA_df_cor_mcf$log2dif > 1 & cDNA_df_cor_mcf$pval < 0.2] <- "Yes"
nutlin_sensitive <- cDNA_df_cor_mcf %>%
  dplyr::select(reporter_id, sensitive)



## Check how well the scaled and normalized reporter activities correlate for the MCF conditions
cDNA_df_cor_mcf <- cDNA_df %>%
  filter(str_detect(sample, "MCF"), pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T)) %>%
  dplyr::select(reporter_id, reporter_activity_condition, condition) %>%
  unique() %>%
  spread(condition, reporter_activity_condition) %>%
  column_to_rownames("reporter_id")


cor_df <- cor(cDNA_df_cor_mcf, method = "pearson", use = "pairwise.complete.obs")

pheatmap(cor_df, border_color = "black")

# Correlation matrix plot
boundaries <- seq(from = 0.8, by = 0.05, length.out = 4)
ggpairs(cDNA_df_cor_mcf,
                 upper = list(continuous = corColor),
               lower = list(continuous = function(data, mapping, ...) {
                   ggally_points(data = data, mapping = mapping, alpha = 0.1, size = 0.5) +
                   geom_abline(slope = 1, lty = "dashed", col = "red") +
                   theme_bw()}),
               diag = list(continuous = function(data, mapping, ...) {
                   ggally_densityDiag(data = data, mapping = mapping, alpha = 0.3, fill = "red") +
                   theme_bw()})) +
  ggtitle(paste("Correlation Between Replicates")) +
  theme(text = element_text(size = 10)) +
  xlab("Reporter activity") +
  ylab("Reporter activity") 


## Investigate more closely
cDNA_df_cor_mcf <- cDNA_df_cor_mcf %>%
  rownames_to_column("reporter_id")
ctrls <- cDNA_df %>%
  dplyr::select(reporter_id, neg_ctrls, tf, promoter, positive_ctrl, motif_id) %>%
  unique()
cDNA_df_cor_mcf <- merge(cDNA_df_cor_mcf, ctrls)

x <- cDNA_df_cor_mcf %>% 
         filter(tf == "P53", motif_id != "4_4_4_4", promoter != "random") %>%
         mutate(dif = MCF7_WT_Nutlin/ MCF7_WT_DMSO) %>%
         na.omit()

x$dif[x$dif > 5] <- 5

ggplot() +
  geom_point(data = x %>% filter(neg_ctrls == "No"), aes(x = MCF7_WT_DMSO, y = MCF7_WT_Nutlin, color = x$dif[x$neg_ctrls == "No"])) +
  scale_color_gradient(low = "#E6EEE8", high = colors_continous[2]) +
  geom_point(data = x %>% filter(neg_ctrls == "Yes"), aes(x = MCF7_WT_DMSO, y = MCF7_WT_Nutlin), color = "black") +
  geom_abline(linetype = "dashed")

## Correlate replicates
cDNA_df_cor_mcf2 <- cDNA_df %>%
  filter(str_detect(sample, "MCF7"), pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T)) %>%
  dplyr::select(reporter_id, reporter_activity_sample, sample) %>%
  unique() %>%
  spread(sample, reporter_activity_sample)
cDNA_df_cor_mcf2 <- merge(cDNA_df_cor_mcf2, ctrls)

ggplot(cDNA_df_cor_mcf2 %>%
         filter(tf == "P53", positive_ctrl == "No"),
       aes(x = MCF7_DMSO_r1_gcf6501, y = MCF7_DMSO_r2_gcf6501, color = neg_ctrls)) +
  geom_point() +
  scale_color_manual(values = c(colors_continous[2], "black"))+
  theme_pubr(border = T)

ggplot(cDNA_df_cor_mcf2 %>%
         filter(tf == "P53", positive_ctrl == "No"),
       aes(x = MCF7_Nutlin_r1_gcf6501, y = MCF7_Nutlin_r2_gcf6501, color = neg_ctrls)) +
  geom_point() +
  scale_color_manual(values = c(colors_continous[2], "black"))+
  theme_pubr(border = T)

## Prove that the KO sample is good enough
cDNA_df_cor_mcf2$ctrl_reporters <- "No"
cDNA_df_cor_mcf2$ctrl_reporters[cDNA_df_cor_mcf2$tf == "GR" & cDNA_df_cor_mcf2$promoter == "mCMV" & cDNA_df_cor_mcf2$neg_ctrls == "Yes"] <- "Yes"

  
ggplot(cDNA_df_cor_mcf2 %>%
           filter(tf == "P53" | ctrl_reporters == "Yes", positive_ctrl == "No"),
       aes(x = MCF7_KO_DMSO_r1_gcf6501, y = MCF7_DMSO_r1_gcf6501, color = ctrl_reporters)) +
  geom_abline(lty = "twodash") +
  geom_point() +
  scale_color_manual(values = c(colors_continous[2], "black")) +
  theme_pubr(border = T)

ggplot(cDNA_df_cor_mcf2 %>%
           filter(tf == "P53" | ctrl_reporters == "Yes", positive_ctrl == "No"),
       aes(x = MCF7_DMSO_r2_gcf6501, y = MCF7_DMSO_r1_gcf6501, color = ctrl_reporters)) +
    geom_point() +
    scale_color_manual(values = c(colors_continous[2], "black"))+
    theme_pubr(border = T)


## Reporter activity per condition, compared to negative and positive controls
data <- cDNA_df %>%
           dplyr::select(reporter_activity_condition, positive_ctrl, promoter, condition, neg_ctrls, outlier, affinity_id, tf, condition, pDNA_library, reporter_id) %>%
           filter(promoter != "random", str_detect(condition, "MCF"), str_detect(condition, "PFT", negate = T), tf == "P53", pDNA_library == "lib_2", 
                  affinity_id != "1_null_only", str_detect(condition, "KO_Nutlin", negate = T)) %>%
           ungroup() %>%
           unique()


ggplot(data, 
       aes(y = reporter_activity_condition, x = condition)) + 
    geom_quasirandom(data = data %>% filter(neg_ctrls == "No"), color = reds[1]) + 
  geom_quasirandom(data = data %>% filter(neg_ctrls == "Yes"), color = reds[3]) +
  #geom_quasirandom(data = data %>% filter(positive_ctrl == "Yes"), color = "black") + 
    xlab("condition") + 
  ylab("reporter activity") +
    theme_pubr(border = T)
```

Conclusion: Replicates do correlate well, there is a strong difference in the degree of Nutlin-specificity.


---

## Figure 2: Reporter framework

Aim: Show that the reporter framework influences reporter activity. Reporter framework = background + promoter (everything except the binding sites)

```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## reporter activity distribution - compare promoters
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No",
                str_detect(condition, "MCF7_WT"), motif_id != "4_4_4_4", tf == "P53", pDNA_library == "lib_2", str_detect(condition, "PFT", negate = T)) %>%
         dplyr::select(condition, reporter_activity_condition, reporter_id, positive_ctrl, promoter, tf) %>%
         unique(),
       aes(x = promoter, y = reporter_activity_condition, color = condition)) + 
  geom_quasirandom(dodge.width = 0.75) +
  geom_boxplot(position = position_dodge(0.75), alpha = 0.7)+
  xlab("") + 
  ylab("reporter activity (a.u.)") +
  labs(title = "reporter activity distribution per minimal promoter") +
  scale_color_manual(values = colors_continous) +
  theme_pubr(border = T)

## reporter activity distribution - enrichment over scrambled per minimal promoter 
x <- cDNA_df %>%
         filter(str_detect(condition, "MCF7_WT"), motif_id != "4_4_4_4", tf == "P53", pDNA_library == "lib_2", str_detect(condition, "PFT", negate = T)) %>%
         dplyr::select(condition, reporter_activity_condition, reporter_id, positive_ctrl, promoter, tf, neg_ctrls) %>%
         unique()


x$reporter_activity_condition[x$promoter == "minP"] <- x$reporter_activity_condition[x$promoter == "minP"] /
  mean(x$reporter_activity_condition[x$promoter == "minP" & x$neg_ctrls == "Yes"])

x$reporter_activity_condition[x$promoter == "mCMV"] <- x$reporter_activity_condition[x$promoter == "mCMV"] /
  mean(x$reporter_activity_condition[x$promoter == "mCMV" & x$neg_ctrls == "Yes"])

x$reporter_activity_condition[x$promoter == "random"] <- x$reporter_activity_condition[x$promoter == "random"] /
  mean(x$reporter_activity_condition[x$promoter == "random" & x$neg_ctrls == "Yes"])

ggplot(x %>% filter(neg_ctrls == "No"),
       aes(x = promoter, y = reporter_activity_condition, color = condition)) + 
  geom_quasirandom(dodge.width = 0.75) +
  geom_boxplot(position = position_dodge(0.75), alpha = 0.7)+
  xlab("") + 
  ylab("reporter activity (a.u.)") +
  labs(title = "reporter activity distribution per minimal promoter") +
  scale_color_manual(values = colors_continous) +
  theme_pubr(border = T)

x <- x %>%
  filter(neg_ctrls == "No") %>%
  mutate(reporter_activity_condition = ave(reporter_activity_condition, promoter, condition, FUN = function(x) mean(x))) %>%
  dplyr::select(reporter_activity_condition, promoter, condition) %>%
  unique()

ggplot(x, aes(x = promoter, y = reporter_activity_condition, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, color = "black")+
  scale_fill_manual(values = colors_continous) +
  theme_pubr(border = T)

## impact of background sequence
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter != "random", 
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), str_detect(affinity_id, "[9]_", negate = T), 
                tf == "P53") %>%
         dplyr::select(tf, reporter_activity_condition, background, reporter_id, condition, affinity_id, promoter) %>% 
         unique(), aes(x = condition, y = reporter_activity_condition, color = factor(background))) +
  geom_quasirandom(dodge.width = .75) +
  geom_boxplot(position = position_dodge(.75), alpha = .8)+
  scale_color_manual(values = c("#000000","#666666", "#999999"))+
  theme_pubr(border = T) +
  labs(title = "reporter activity per background")+
  xlab("background sequence")+
  ylab("reporter activity")+
  facet_wrap(~promoter, scales = "free")+ 
  stat_compare_means(method = "t.test") 

## correlate minP and mCMV reporter activities
data <- cDNA_df %>%
  filter(neg_ctrls == "No", positive_ctrl == "No", promoter != "random", motif_id != "4_4_4_4",
         str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), 
         tf == "P53", pDNA_library == "lib_2") %>%
  dplyr::select(tf, reporter_activity_condition, n_sites, motif_id, condition, reporter_id_2, background, promoter) %>% 
  unique() %>%
  spread(promoter, reporter_activity_condition) %>% 
  na.omit()

# reporters <- sample(data$reporter_id_2, 300)
# data <- data[data$reporter_id_2 %in% reporters,]

ggplot(data, aes(x = minP, y = mCMV)) +
  geom_point(size = 1, aes(color = factor(background))) +
  geom_smooth(method = "lm", aes(color = factor(background), fill = factor(background))) +
  geom_abline(slope = mean(data$mCMV)/mean(data$minP), linetype = "dashed") +
  theme_pubr(border = T) +
  scale_color_manual(values = c("#000000","#666666", "#999999")) +
  scale_fill_manual(values = c("#000000","#666666", "#999999"))


## minimal promoters background activity
data <- cDNA_df %>%
  filter(neg_ctrls == "Yes", positive_ctrl == "No",
         str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), 
         tf == "P53", pDNA_library == "lib_2") %>%
  dplyr::select(tf, reporter_activity_condition, n_sites, motif_id, condition, reporter_id_2, background, promoter) %>% 
  unique() %>%
  na.omit()

ggplot(data, aes(x = promoter, y = reporter_activity_condition)) +
  geom_quasirandom() +
  geom_boxplot(aes(fill = promoter)) +
  scale_fill_manual(values = c("#F6D289", "#ECD9B1", "#FBEDD0"))+
  theme_pubr(border = T)

# reporters with 0bp spacing
data <- cDNA_df %>%
  filter(neg_ctrls == "No", positive_ctrl == "No", promoter != "random", motif_id != "4_4_4_4", spacing == 0,
         str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), 
         tf == "P53", pDNA_library == "lib_2") %>%
  dplyr::select(tf, reporter_activity_condition, n_sites, motif_id, condition, reporter_id_2, background, promoter) %>% 
  unique()%>%
  spread(promoter, reporter_activity_condition) %>% 
  na.omit()

ggplot(data, aes(x = minP, y = mCMV)) +
  geom_point(aes(color = factor(background))) +
  geom_smooth(se = F, method = "lm", aes(color = factor(background), fill = factor(background))) +
  geom_abline(slope = mean(data$mCMV)/mean(data$minP), linetype = "dashed") +
  theme_pubr(border = T) +
  scale_color_manual(values = c("#000000","#666666", "#999999")) +
  scale_fill_manual(values = c("#000000","#666666", "#999999"))
```

Conclusion: mCMV is the strongest, minimal promoter interacts with upstream background sequences.

---

Figure 3: Binding site affinity and spacing

Aim: Show how P53 affinity and binding site spacing affect reporter activity. 

*ADD: FOLLOW-UP LIBRARY AFFINITY*

```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## activity per spacing at different affinities - lib 2 all data combined
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", promoter == "mCMV", positive_ctrl == "No", pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T),
                str_detect(condition, "MCF7_WT"), str_detect(affinity_id, "[1,3,5,7]{1}_"), str_detect(affinity_id, "11_high_mid", negate = T),
                position == 0, tf == "P53") %>%
         dplyr::select(tf, reporter_activity_condition, spacing, reporter_id, condition, affinity_id) %>% 
         unique() %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, affinity_id, condition, spacing, FUN = function(x) mean(x))) %>%
         unique(), 
       aes(x = spacing, y = reporter_activity_condition, color = condition, fill = condition)) +
  geom_point() +
  geom_smooth()+
  theme_pubr(border = T) +
  ggtitle("reporter activity per spacing")+
  xlab("spacing length between motifs (bp)")+
  ylab("reporter activity (a.u.)")+
  scale_color_manual(values = colors_continous)+
  scale_fill_manual(values = colors_continous)+
  facet_wrap(~affinity_id, nrow = 1) +
  scale_x_continuous(breaks = seq(0,10,1))


## activity per spacing at different affinities - only low_affinity, background 3
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", promoter != "random", positive_ctrl == "No", pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T),
                str_detect(condition, "MCF7_WT"), str_detect(affinity_id, "[7]_"), position == 0, tf == "P53") %>%
         dplyr::select(tf, reporter_activity_condition, spacing, reporter_id, condition, affinity_id, background, promoter) %>% 
         unique() %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, affinity_id, condition, background, promoter, spacing, FUN = function(x) mean(x))) %>%
         unique(), 
       aes(x = spacing, y = reporter_activity_condition, color = condition, fill = condition)) +
  geom_point() +
  geom_smooth(se = F)+
  theme_pubr(border = T) +
  ggtitle("reporter activity per spacing")+
  xlab("spacing length between motifs (bp)")+
  ylab("reporter activity (a.u.)")+
  scale_color_manual(values = colors_continous)+
  scale_fill_manual(values = colors_continous)+
  facet_grid(promoter~background, scales = "free") +
  scale_x_continuous(breaks = seq(0,10,1))


## activity per spacing - mean over all affinities
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", promoter == "mCMV", positive_ctrl == "No", pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T),
                str_detect(condition, "MCF7_WT"), str_detect(affinity_id, "[1,3,5,7]_"), position == 0, tf == "P53") %>%
         dplyr::select(tf, reporter_activity_condition, spacing, condition, background, affinity_id) %>% 
         unique() %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, affinity_id, condition, background, spacing, FUN = function(x) mean(x))) %>%
         unique(), 
       aes(x = spacing, y = reporter_activity_condition, color = condition, fill = condition)) +
  geom_quasirandom() +
  geom_smooth(method = "loess")+
  stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 5, label.y = 8) +
  theme_pubr(border = T) +
  ggtitle("reporter activity per spacing")+
  xlab("spacing length between motifs (bp)")+
  ylab("reporter activity (a.u.)")+
  scale_color_manual(values = rev(reds[seq(1, length(reds), 3)]))+
  scale_fill_manual(values = rev(reds[seq(1, length(reds), 3)]))+
  scale_x_continuous(breaks = seq(0,10,1))


## activity per affinity
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", promoter == "mCMV", positive_ctrl == "No", spacing == 7, tf == "P53",
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), 
                str_detect(affinity_id, "[1,3,5,7,9]_"), str_detect(affinity_id, "11_high_mid", negate = T), pDNA_library == "lib_2") %>%
         dplyr::select(tf, reporter_activity_condition, background, reporter_id, condition, affinity_id) %>%
         unique(), 
       aes(x = affinity_id, y = reporter_activity_condition)) +
  geom_boxplot(position = position_dodge(.75), aes(fill = condition), alpha = 0.7) +
  geom_quasirandom(dodge.width = .75, aes(color = affinity_id, group = condition)) +
  theme_pubr(border = T) +
  xlab("") +
  scale_color_manual(values = reds) +
  scale_fill_manual(values = colors_continous[c(1,2)]) +
  ggtitle("activity per affinity (P53-mCMV only)") + 
  stat_compare_means(method = "t.test") 

## DMSO vs Nutlin depending on reporter affinity
x <- cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "minP",
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T),
                tf == "P53", spacing == 7, background == 1, position == 0, n_sites == 4, str_detect(affinity_id, "[1,2,3,4,5,6,7]{1}_"),
                str_detect(affinity_id, "10_high_start|11_high_mid|12_high_end", negate = T)) %>%
         dplyr::select(reporter_activity_condition, condition, SumAffinity, reporter_id, affinity_id) %>%
         unique() %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, condition, SumAffinity,
                                                  FUN = function(x) mean(x))) %>%
         unique() %>%
         arrange(SumAffinity) %>%
         group_by(condition) %>%
         mutate(SumAffinity = 1:length(SumAffinity)) %>%
         spread(condition, reporter_activity_condition) %>%
         mutate(dif = MCF7_WT_Nutlin / MCF7_WT_DMSO)

x$copies <- "mixed"
x$copies[grep("[1,3,5,7]_", x$affinity_id)] <- "non_mixed"

ggplot(x,
       aes(x = MCF7_WT_DMSO, y = MCF7_WT_Nutlin, color = SumAffinity, shape = copies)) +
  geom_point(size = 2) +
  scale_color_gradient2(low = "grey", mid = "#F9E9E4", high = reds[1], midpoint = 24) +
  geom_abline(linetype = "dashed")+
  theme_pubr(border = T) +
  labs(title = "reporter activity per affinity")

## Rule of thumb for fitting models: e^n_features < n_observations -> I will try to fit 4 features for 114 observations

## Subselect reporters
cDNA_df_p53 <- cDNA_df %>%
  filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", tf == "P53",
         position == 0, promoter != "random",
         str_detect(affinity_id, "[1,3,5,7]_"), str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T)) %>%
  dplyr::select(reporter_id, condition, reporter_activity_condition, 
                affinity_pos1, spacing, background, promoter) %>%
  mutate(spacing = as.factor(spacing),
         background = as.factor(background),
         reporter_activity = ave(reporter_activity_condition, affinity_pos1, background, promoter, condition, spacing, FUN = function(x) mean(x)),
         log_reporter_activity = log2(reporter_activity)) %>%
  dplyr::select(-reporter_activity_condition, -reporter_activity) %>%
  unique() 


# Rename 0,1,2,3,4 in null, very-weak, weak, medium, strong
replace <- data.frame("old" = c(0:4), "new" = c("motif 1", "motif 2", "motif 3", "motif 4", "motif 0"), stringsAsFactors=FALSE)
for (i in unique(replace$old)) {
  cDNA_df_p53$affinity_pos1[cDNA_df_p53$affinity_pos1 == i] <- replace$new[replace$old == i]
}



## First for DMSO
cDNA_df_p53_dmso <- cDNA_df_p53[,-1] %>% filter(cDNA_df_p53$condition == "MCF7_WT_DMSO")
cDNA_df_p53_dmso$row <- rownames(cDNA_df_p53_dmso)

## Fitting the model
x <- lm(log_reporter_activity ~ affinity_pos1 + spacing + background:promoter, cDNA_df_p53_dmso)

summ(x)
par(mfrow=c(2,2))
plot(x)

# Get predicted activities per reporter
y <- data.frame(x$fitted.values, stringsAsFactors=FALSE) %>% rownames_to_column()
names(y) <- c("row", "reporter_activity_predicted")
y$id <- 1:nrow(y)
prediction <- data.frame("id" = 1:(nrow(cDNA_df_p53_dmso)), stringsAsFactors=FALSE)
prediction <- merge(prediction,y, all = T) %>% 
  dplyr::select(-id)
cDNA_df_p53_dmso <- merge(cDNA_df_p53_dmso, prediction)
  

# What are the individual weights?
weight <- data.frame(x$coefficients, stringsAsFactors=FALSE) %>% rownames_to_column() %>% na.omit()
names(weight) <- c("feature", "weight")
references <- data.frame("feature" = c("affinity_pos1motif1", "spacing0", "background3:promoterminP"), 
                         "weight" = c(0,0,0))
weight <- rbind(weight, references)
weight <- weight %>% filter(feature != "(Intercept)")
categorie <- c("affinity_pos1", "spacing")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
weight$cond <- gsub(paste(weight$features, collapse = "|"), "", weight$feature)
weight$cond[weight$cond == ""] <- "background:promoter"
weight$condition <- "DMSO"


## Now for Nutlin
cDNA_df_p53_nutlin <- cDNA_df_p53[,-1] %>% filter(cDNA_df_p53$condition == "MCF7_WT_Nutlin")
cDNA_df_p53_nutlin$row <- rownames(cDNA_df_p53_nutlin)

## Fitting the model
x <- lm(log_reporter_activity ~ affinity_pos1 + spacing + background:promoter, cDNA_df_p53_nutlin)

summ(x)
par(mfrow=c(2,2))
plot(x)

# Get predicted activities per reporter
y <- data.frame(x$fitted.values, stringsAsFactors=FALSE) %>% rownames_to_column()
names(y) <- c("row", "reporter_activity_predicted")
y$id <- 1:nrow(y)
prediction <- data.frame("id" = 1:(nrow(cDNA_df_p53_nutlin)), stringsAsFactors=FALSE)
prediction <- merge(prediction,y, all = T) %>% 
  dplyr::select(-id)
cDNA_df_p53_nutlin <- merge(cDNA_df_p53_nutlin, prediction)
cDNA_df_p53 <- rbind(cDNA_df_p53_nutlin, cDNA_df_p53_dmso)
  

# What are the individual weights?
weight_nutlin <- data.frame(x$coefficients, stringsAsFactors=FALSE) %>% rownames_to_column() %>% na.omit()
names(weight_nutlin) <- c("feature", "weight")
references <- data.frame("feature" = c("affinity_pos1motif1", "spacing0", "background3:promoterminP"), 
                         "weight" = c(0,0,0))
weight_nutlin <- rbind(weight_nutlin, references)
weight_nutlin <- weight_nutlin %>% filter(feature != "(Intercept)")
categorie <- c("affinity_pos1", "spacing")
weight_nutlin$features <- gsub(paste(categorie, collapse="|"), "",weight_nutlin$feature)
weight_nutlin$cond <- gsub(paste(weight_nutlin$features, collapse = "|"), "", weight_nutlin$feature)
weight_nutlin$cond[weight_nutlin$cond == ""] <- "background:promoter"
weight_nutlin$condition <- "Nutlin"
weight <- rbind(weight, weight_nutlin)



# Plotting
ggscatter(cDNA_df_p53, x = "log_reporter_activity", y = "reporter_activity_predicted", color = "condition", shape = "promoter",
          add = "reg.line",
          add.params = list(color = "#84A98C", fill = "lightgray"),
          conf.int = TRUE, ylab = "Predicted reporter activity",
          xlab = "Average reporter activity") + 
  theme_pubr(border = T)+
  scale_color_manual(values = colors_continous)

level_order <- c("affinity_pos1motif1", "affinity_pos1motif 2", "affinity_pos1motif 3", "affinity_pos1motif 4", "spacing0", "spacing1", "spacing2", 
                 "spacing3", "spacing4", "spacing5", "spacing6", "spacing7", "spacing8", "spacing9", "spacing10", "background1:promotermCMV", "background2:promotermCMV",
                 "background3:promotermCMV", "background1:promoterminP", "background2:promoterminP", "background3:promoterminP")

ggplot(weight, 
       aes(x = factor(feature, level = level_order), y = weight, fill = cond, group = condition)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") + 
  ylab("weight - reporter activity") +
  xlab("") +
  theme_pubr(border = T) +
  scale_fill_manual(values = colors_continous_2[c(1,3,5,6)]) +
  guides(fill = F)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))



## Show that 4 copies + 7 bp spacing is not the ideal set
cDNA_df_cor_mcf <- cDNA_df %>%
  filter(str_detect(sample, "MCF"), str_detect(sample, "KO", negate = T), pDNA_library == "lib_2", str_detect(sample, "PFT", negate = T),
         tf == "P53", neg_ctrls == "No", positive_ctrl == "No") %>%
  dplyr::select(reporter_id, reporter_activity_condition, condition, spacing, n_sites, promoter, motif_id) %>%
  unique()

cDNA_df_cor_mcf$design <- "alt"
cDNA_df_cor_mcf$design[cDNA_df_cor_mcf$n_sites == 4 & cDNA_df_cor_mcf$spacing == 7] <- "common"

ggplot(cDNA_df_cor_mcf %>% 
         filter(motif_id != "4_4_4_4", promoter == "mCMV") %>%
         na.omit(), 
       aes(x = design, y = reporter_activity_condition, fill = condition)) +
  geom_violin(position = position_dodge(0.75)) +
  theme_pubr(border = T) +
  #scale_color_viridis_b(breaks = c(1,3,6), begin = 0.2, option = "inferno")+
  scale_fill_manual(values = colors_continous)
```

Conclusion:

---

## Figure 4: Positioning of binding sites

Aim: Show how binding site positioning affects reporter activity.

```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## activity for different number of binding sites
data <- cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", promoter == "mCMV", 
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), str_detect(affinity_id, "[7-9]_"), 
                tf == "P53", spacing == 7, pDNA_library == "lib_2") %>%
         dplyr::select(tf, reporter_activity_condition, n_sites, motif_id, condition, affinity_id, gcf) %>% 
         unique() %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x))) %>%
         unique()

ggplot(data, 
       aes(x = n_sites, y = reporter_activity_condition, color = condition)) +
  geom_point(position = position_dodge(0.75))+
  theme_pubr(border = T) +
  geom_label(data = data %>% 
               filter(n_sites != 0, n_sites != 4) %>%
               dplyr::select(reporter_activity_condition, motif_id, condition, n_sites) %>%
               arrange(desc(reporter_activity_condition)) %>% 
               group_by(n_sites) %>% 
               slice_head(n=3), 
             aes(label = motif_id), 
             nudge_x = -.3) +
  ggtitle("reporter activity vs number of binding sites")+
  xlab("number of low-affinity binding sites")+
  ylab("reporter activity (a.u.)")+
  scale_color_manual(values = colors_continous)

## bar plot as before - but only fair comparisons (remove the positioning effect)
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "mCMV", 
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), str_detect(motif_id, "4_4_4_4|4_4_4_3|4_4_3_3|4_3_3_3|3_3_3_3"), 
                tf == "P53", spacing == 7) %>%
         dplyr::select(tf, reporter_activity_condition, n_sites, condition, reporter_id, promoter) %>% 
         unique() %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, n_sites, promoter, condition, FUN = function(x) mean(x))) %>%
         unique(), 
       aes(x = n_sites, y = reporter_activity_condition, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, color = "black")+
  theme_pubr(border = T) +
  ggtitle("reporter activity vs number of binding sites")+
  xlab("number of low-affinity binding sites")+
  ylab("reporter activity (a.u.)")+
  scale_fill_manual(values = colors_continous)


## compare all reporters with at least one high-affinity binding site
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "mCMV", position == 0, 
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), spacing == 7,
                tf == "P53", str_detect(motif_id, "0_0_0_0|0_4_4_0|4_4_0_0|0_0_4_4|4_0_0_4")) %>%
         dplyr::select(reporter_activity_condition, condition, motif_id) %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, condition, motif_id, FUN = function(x) mean(x))) %>%
         unique(),
       aes(x = motif_id, y = reporter_activity_condition, fill = condition))+
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, color = "black")+
  theme_pubr(border = T) +
  ylab("reporter activity (a.u.)")+
  xlab("motif ID")+
  scale_fill_manual(values = colors_continous)


## disruption of activity by adding high-affinity binding sites
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "mCMV", position == 0, spacing == 7, 
                str_detect(condition, "MCF7_WT_Nutlin"), str_detect(condition, "PFT", negate = T),
                tf == "P53", str_detect(motif_id, "3_3_3_3|0_0_3_3|4_4_3_3|0_0_4_4|3_3_4_4|4_4_4_4|4_4_0_0|3_3_0_0|0_0_0_0|0_3_3_0|0_4_4_0|3_4_4_3")) %>%
         dplyr::select(reporter_activity_condition, condition, motif_id) %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x))) %>%
         unique(),
       aes(x = motif_id, y = reporter_activity_condition, fill = condition))+
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, color = "black")+
  theme_pubr(border = T) +
  ylab("reporter activity (a.u.)")+
  xlab("motif ID") +
  scale_x_discrete(limits = rev) +
  scale_fill_manual(values = colors_continous[2])

## disruption of activity by adding high-affinity binding sites
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "minP", position == 0, spacing == 7, 
                str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T),
                tf == "P53", str_detect(motif_id, "3_3_3_3|0_0_3_3|3_3_0_0|0_0_0_0|0_3_3_0")) %>%
         dplyr::select(reporter_activity_condition, condition, motif_id) %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x))) %>%
         unique(),
       aes(x = motif_id, y = reporter_activity_condition, fill = condition))+
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, color = "black")+
  theme_pubr(border = T) +
  ylab("reporter activity (a.u.)")+
  xlab("motif ID") +
  scale_x_discrete(limits = rev) +
  scale_fill_manual(values = colors_continous)

## Is the activity lower when the higher affinity binding site is further from the TSS?
x <- cDNA_df %>%
         filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "mCMV", position == 0, spacing == 7,
                str_detect(condition, "MCF7_WT_Nutlin"), str_detect(condition, "PFT", negate = T),
                tf == "P53", str_detect(motif_id, "2_3_3_3|3_3_3_2|1_1_1_0|0_1_1_1|2_2_2_1|1_2_2_2|2_2_3_3|3_3_2_2|1_1_2_2|2_2_1_1|1_1_0_0|0_0_1_1")) %>%
         dplyr::select(reporter_activity_condition, condition, reporter_id, motif_id, affinity_id, SumAffinity) %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x)),
                SumAffinity = round(SumAffinity, 2)) %>%
         unique()
x$positioning <- "proximal"
x$positioning[grep("2_3_3_3|0_1_1_1|1_2_2_2|2_2_3_3|1_1_2_2|0_0_1_1", x$motif_id)] <- "distal"
x <- x %>%
  dplyr::select(-motif_id, -reporter_id) %>%
  unique()

ggplot(x,
       aes(x = positioning, y = reporter_activity_condition))+
  geom_boxplot(alpha = 0.4)+
  geom_abline()+
  theme_pubr(border = T)+ 
  stat_compare_means() 

## Can I quantify this somehow else?
cDNA_df_2 <- cDNA_df
cDNA_df_2$positioning <- "other"
cDNA_df_2$positioning[grep("2_3_3_3|0_1_1_1|1_2_2_2|2_2_3_3|0_0_1_1|1_1_2_2", cDNA_df_2$motif_id)] <- "high_end"
cDNA_df_2$positioning[grep("3_3_3_2|1_1_1_0|2_2_2_1|3_3_2_2|1_1_0_0|2_2_1_1", cDNA_df_2$motif_id)] <- "high_start"
cDNA_df_2$motif_id2 <- "other"
cDNA_df_2$motif_id2[grep("2_3_3_3|3_3_3_2", cDNA_df_2$motif_id)] <- "2_one"
cDNA_df_2$motif_id2[grep("0_1_1_1|1_1_1_0", cDNA_df_2$motif_id)] <- "0_one"
cDNA_df_2$motif_id2[grep("1_2_2_2|2_2_2_1", cDNA_df_2$motif_id)] <- "1_one"
cDNA_df_2$motif_id2[grep("2_2_3_3|3_3_2_2", cDNA_df_2$motif_id)] <- "2_two"
cDNA_df_2$motif_id2[grep("0_0_1_1|1_1_0_0", cDNA_df_2$motif_id)] <- "0_two"
cDNA_df_2$motif_id2[grep("1_1_2_2|2_2_1_1", cDNA_df_2$motif_id)] <- "1_two"

ggplot(cDNA_df_2 %>%
          filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", promoter == "mCMV", position == 0, spacing == 7,
                str_detect(condition, "MCF7_WT_Nutlin"), str_detect(condition, "PFT", negate = T),
                tf == "P53", str_detect(positioning, "high")) %>%
         dplyr::select(reporter_activity_condition, positioning, motif_id2, background) %>%
         mutate(reporter_activity_condition = ave(reporter_activity_condition, motif_id2, background, positioning, FUN = function(x) mean(x))) %>%
         unique() %>%
         spread(positioning, reporter_activity_condition) %>%
         mutate(dif = high_start - high_end),
       aes(x = high_end, y = high_start, color = dif))+ 
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(linetype = "dashed") +
  theme_pubr(border = T) +
  scale_color_viridis_b(option = "inferno") +
  xlim(50, 210) +
  ylim(50, 210)

## Rule of thumb for fitting models: e^n_features < n_observations

## Subselect p53 only 
cDNA_df_p53 <- cDNA_df %>%
  filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", tf == "P53",
         spacing == 7, str_detect(affinity_id, "[7-9]"), position == 0, 
         promoter == "mCMV", str_detect(condition, "MCF7_WT_Nutlin"), str_detect(condition, "PFT", negate = T)) %>%
  dplyr::select(reporter_id, condition, reporter_activity_condition,
                affinity_pos1, affinity_pos2, affinity_pos3, affinity_pos4, motif_id, n_sites) %>% 
  mutate(reporter_activity = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x)),
         log_reporter_activity = log2(reporter_activity)) %>%
  dplyr::select(-reporter_id, - reporter_activity_condition) %>%
  unique()

# cDNA_df_p53$motif_dist <- 0
# cDNA_df_p53$motif_dist[grep("3_4_4$", cDNA_df_p53$motif_id)] <- 1
# cDNA_df_p53$motif_dist[grep("3_4$", cDNA_df_p53$motif_id)] <- 2
# cDNA_df_p53$motif_dist[grep("3$", cDNA_df_p53$motif_id)] <- 3

cDNA_df_p53$highest_dist <- "0"
cDNA_df_p53$highest_dist[grep("3_4_3", cDNA_df_p53$motif_id)] <- "1"
cDNA_df_p53$highest_dist[grep("3_4_4_3", cDNA_df_p53$motif_id)] <- "1"

#cDNA_df_p53$n_adjacent <- "0"
#cDNA_df_p53$n_adjacent[grep("3_3", cDNA_df_p53$motif_id)] <- "2"
#cDNA_df_p53$n_adjacent[grep("3_3_3", cDNA_df_p53$motif_id)] <- "3"

#cDNA_df_p53$n_sites <- as.character(cDNA_df_p53$n_sites)

# Rename 0,1,2,3,4 in null, very-weak, weak, medium, strong
replace <- data.frame("old" = c(0:4), "new" = c("_4_strong", "_3_medium", "_2_weak", "_1_very-weak", "_0_null"), stringsAsFactors=FALSE)
for (i in unique(replace$old)) {
  cDNA_df_p53$affinity_pos1[cDNA_df_p53$affinity_pos1 == i] <- replace$new[replace$old == i]
  cDNA_df_p53$affinity_pos2[cDNA_df_p53$affinity_pos2 == i] <- replace$new[replace$old == i]
  cDNA_df_p53$affinity_pos3[cDNA_df_p53$affinity_pos3 == i] <- replace$new[replace$old == i]
  cDNA_df_p53$affinity_pos4[cDNA_df_p53$affinity_pos4 == i] <- replace$new[replace$old == i]
}

## Explain which affinities are essential at which position to boost activity
cDNA_df_p53 <- cDNA_df_p53[order(cDNA_df_p53$condition),] %>%
  dplyr::select(-motif_id)
cDNA_df_p53$row <- rownames(cDNA_df_p53)

## Fitting the model
x <- lm(reporter_activity ~ n_sites + highest_dist + affinity_pos4, cDNA_df_p53)

summ(x)
par(mfrow=c(2,2))
plot(x, main = i)

# Get predicted activities per reporter
y <- data.frame(x$fitted.values, stringsAsFactors=FALSE) %>% rownames_to_column()
names(y) <- c("row", "reporter_activity_predicted")
y$id <- 1:nrow(y)
prediction <- data.frame("id" = 1:(nrow(cDNA_df_p53)), stringsAsFactors=FALSE)
prediction <- merge(prediction,y, all = T) %>% 
  dplyr::select(-id)
cDNA_df_p53 <- merge(cDNA_df_p53, prediction)
  

# What are the individual weights?
weight <- data.frame(x$coefficients, stringsAsFactors=FALSE) %>% rownames_to_column()
names(weight) <- c("feature", "weight")
weight <- weight %>% filter(feature != "(Intercept)")
categorie <- c("affinity_pos", "n_sites", "highest_dist")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
weight$cond <- gsub(paste(weight$features, collapse = "|"), "", weight$feature)
  

# How many percent do the different tested features explain?
exp_variance <- anova(x) %>% rownames_to_column("feature") %>%
  setnames("Sum Sq", "sum_sq") %>%
  dplyr::select(feature, sum_sq) %>%
  mutate(sum = sum(sum_sq)) %>%
  mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
  setnames("rel_sum_sq", "percent") %>%
  dplyr::select("feature", "percent")

exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
exp_variance$percent[exp_variance$feature == "Residuals"] <- 100 - exp_variance$percent[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))


# Plotting
ggplot(exp_variance, 
       aes(x = percent, y = reorder(feature, -percent), fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  xlab("Variance explained (%)") + 
  ylab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_pubr(border = T) +
  theme(text = element_text(size = 12)) +
  guides(fill = F)

ggscatter(cDNA_df_p53, x = "reporter_activity", y = "reporter_activity_predicted",
          add = "reg.line",
          add.params = list(color = "#84A98C", fill = "lightgray"),
          conf.int = TRUE, ylab = "Predicted reporter activity", size = 3,
          xlab = "Average reporter activity") + 
  theme_pubr(border = T)+
  stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 0, label.y = 3)

ggplot(weight, 
       aes(x = reorder(feature, -weight), y = weight)) +
  geom_bar(stat = "identity", width = 0.7, fill = colors_continous[2], color = "black") + 
  ylab("weight - reporter activity") +
  xlab("") +
  theme_pubr(border = T) +
  guides(fill = F)
```

Conclusion: 

---

## Figure 5: Identification of super-reporters

Aim: Can we show that with this systematic screen we were able to identify reporters that are superior (more sensitive & specific) to the commonly used P53 reporters?

*ADD: FOLLOW-UP SUPERIOR REPORTERS*

```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# Plotting specificity
x <- cDNA_df_cor_mcf %>% 
         filter(tf == "P53", motif_id != "4_4_4_4", promoter != "random") %>%
         mutate(dif = MCF7_WT_Nutlin / MCF7_WT_DMSO) %>%
         na.omit()

ggplot(x,
       aes(x = MCF7_WT_DMSO, y = dif, color = positive_ctrl)) +
  geom_point()+
  theme_pubr(border = T) +
  scale_color_manual(values = c("black", "red")) +
  geom_hline(yintercept = 5, linetype = "dashed")

nutlin_specific <- x[x$dif >= 5 | x$positive_ctrl == "Yes",]

ggplot(nutlin_specific,
       aes(x = MCF7_WT_DMSO, y = dif, color = positive_ctrl)) +
  geom_point()+
  theme_pubr(border = T)


x <- cDNA_df_cor_mcf %>% 
         filter(tf == "P53", motif_id != "4_4_4_4", promoter != "random", neg_ctrls == "No") %>%
         mutate(dif = MCF7_WT_DMSO / MCF7_KO_DMSO) %>%
         na.omit()

x$dmso_specific <- "No"
x$dmso_specific[x$dif >= 30] <- "Yes"
x <- x %>%
  filter(MCF7_KO_DMSO >= 0.05)

ggplot(x,
       aes(x = MCF7_WT_DMSO, y = dif, color = positive_ctrl)) +
  geom_point()+
  scale_color_manual(values = c("black", "red")) +
  geom_hline(yintercept = 30, linetype = "dashed") +
  theme_pubr(border = T)


### Super-reporter?
super_reporter <- cDNA_df %>%
  filter(str_detect(reporter_id,
                    "P53_mCMV_p_0_s_7_d_10_bg_1_a1_3_a2_4_a3_3_a4_4|P53_minP_p_0_s_7_d_10_bg_3_a1_3_a2_4_a3_3_a4_4|P53_mCMV_p_0_s_1_d_10_bg_1_a1_2_a2_2_a3_2_a4_2|P53_minP_p_0_s_1_d_10_bg_3_a1_2_a2_2_a3_2_a4_2|P53_minP_p_70_s_7_d_10_bg_3_a1_0_a2_0_a3_0_a4_0|P53_mCMV_p_70_s_7_d_10_bg_1_a1_0_a2_0_a3_0_a4_0"),
         str_detect(condition, "MCF7"),
         str_detect(condition, "PFT", negate = T),
         pDNA_library == "lib_2",
         promoter != "random") %>%
  dplyr::select(condition, reporter_activity_condition, promoter, background, positive_ctrl, motif_id) %>%
  unique()

ggplot(super_reporter, aes(x = motif_id, y = reporter_activity_condition, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, color = "black") +
  theme_pubr(border = T) +
  scale_fill_manual(values = colors_continous) +
  facet_wrap(~promoter, scales = "free")
```

Conlusion: 























<!-- --- Testing of other linear & machine learning models --- -->


<!-- ### Explain expression differences betweeen all other mixed populations - how does high affinity binding influence activity? -->
<!-- ```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE} -->
<!-- ## Rule of thumb for fitting models: e^n_features < n_observations -->

<!-- ## Subselect p53 only - let's only go for 2-2 combinations for now to simplify the model -->
<!-- cDNA_df_p53 <- cDNA_df %>% -->
<!--   filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", tf == "P53", -->
<!--          spacing == 7, str_detect(affinity_id, "[2,4,6]_"), position == 0, background == 1, n_sites == 4, -->
<!--          promoter == "mCMV", str_detect(condition, "MCF7_WT_Nutlin"), str_detect(condition, "PFT", negate = T)) %>% -->
<!--   dplyr::select(reporter_id, condition, reporter_activity_condition, -->
<!--                 affinity_pos1, affinity_pos2, affinity_pos3, affinity_pos4, motif_id, n_sites, affinity_id, -->
<!--                 SumAffinity) %>% -->
<!--   mutate(reporter_activity = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x)), -->
<!--          log_reporter_activity = log2(reporter_activity), -->
<!--          SumAffinity = round(SumAffinity, 2), -->
<!--          SumAffinity = as.character(SumAffinity)) %>% -->
<!--   dplyr::select(-reporter_id, - reporter_activity_condition) %>% -->
<!--   unique() -->

<!-- cDNA_df_p53 <- cDNA_df_p53 %>% -->
<!--          filter(str_detect(motif_id, "2_3_3_3|3_3_3_2|1_1_1_0|0_1_1_1|2_2_2_1|1_2_2_2|2_2_3_3|3_3_2_2|1_1_2_2|2_2_1_1|1_1_0_0|0_0_1_1")) -->
<!-- cDNA_df_p53$positioning <- "proximal" -->
<!-- cDNA_df_p53$positioning[grep("2_3_3_3|0_1_1_1|1_2_2_2|2_2_3_3|1_1_2_2|0_0_1_1", cDNA_df_p53$motif_id)] <- "distal" -->

<!-- ## Explain which affinities are essential at which position to boost activity -->
<!-- cDNA_df_p53 <- cDNA_df_p53[order(cDNA_df_p53$condition),] %>% -->
<!--   dplyr::select(-motif_id) -->
<!-- cDNA_df_p53$row <- rownames(cDNA_df_p53) -->

<!-- ## Fitting the model -->
<!-- x <- lm(reporter_activity ~ SumAffinity + positioning, cDNA_df_p53) -->

<!-- summ(x) -->
<!-- par(mfrow=c(2,2)) -->
<!-- plot(x, main = i) -->

<!-- # Get predicted activities per reporter -->
<!-- y <- data.frame(x$fitted.values, stringsAsFactors=FALSE) %>% rownames_to_column() -->
<!-- names(y) <- c("row", "reporter_activity_predicted") -->
<!-- y$id <- 1:nrow(y) -->
<!-- prediction <- data.frame("id" = 1:(nrow(cDNA_df_p53)), stringsAsFactors=FALSE) -->
<!-- prediction <- merge(prediction,y, all = T) %>% -->
<!--   dplyr::select(-id) -->
<!-- cDNA_df_p53 <- merge(cDNA_df_p53, prediction) -->


<!-- # What are the individual weights? -->
<!-- weight <- data.frame(x$coefficients, stringsAsFactors=FALSE) %>% rownames_to_column() -->
<!-- names(weight) <- c("feature", "weight") -->
<!-- weight <- weight %>% filter(feature != "(Intercept)") -->
<!-- categorie <- c("affinity_pos", "n_sites", "highest_dist") -->
<!-- weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature) -->
<!-- weight$cond <- gsub(paste(weight$features, collapse = "|"), "", weight$feature) -->


<!-- # How many percent do the different tested features explain? -->
<!-- exp_variance <- anova(x) %>% rownames_to_column("feature") %>% -->
<!--   setnames("Sum Sq", "sum_sq") %>% -->
<!--   dplyr::select(feature, sum_sq) %>% -->
<!--   mutate(sum = sum(sum_sq)) %>% -->
<!--   mutate(rel_sum_sq = (sum_sq/sum)*100) %>% -->
<!--   setnames("rel_sum_sq", "percent") %>% -->
<!--   dplyr::select("feature", "percent") -->

<!-- exp_variance$group <- "no residual" -->
<!-- exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual" -->
<!-- exp_variance$percent[exp_variance$feature == "Residuals"] <- 100 - exp_variance$percent[exp_variance$feature == "Residuals"] -->
<!-- exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total")) -->


<!-- # Plotting -->
<!-- ggplot(exp_variance, -->
<!--        aes(x = percent, y = reorder(feature, -percent), fill = group)) + -->
<!--   scale_fill_manual(values = c("#1B998B", "#2D3047")) + -->
<!--   geom_bar(stat = "identity") + -->
<!--   xlab("Variance explained (%)") + -->
<!--   ylab("") + -->
<!--   labs(title = "Linear variance modelling", -->
<!--        subtitle = "Expression variance predicted by features") + -->
<!--   theme_pubr(border = T) + -->
<!--   theme(text = element_text(size = 12)) + -->
<!--   guides(fill = F) -->

<!-- ggscatter(cDNA_df_p53, x = "reporter_activity", y = "reporter_activity_predicted", -->
<!--           add = "reg.line", -->
<!--           add.params = list(color = "#84A98C", fill = "lightgray"), -->
<!--           conf.int = TRUE, ylab = "Predicted reporter activity", size = 3, -->
<!--           xlab = "Average reporter activity") + -->
<!--   theme_pubr(border = T)+ -->
<!--   stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 0, label.y = 3) -->

<!-- ggplot(weight, -->
<!--        aes(x = reorder(feature, -weight), y = weight)) + -->
<!--   geom_bar(stat = "identity", width = 0.7, fill = colors_continous[2], color = "black") + -->
<!--   ylab("weight - reporter activity") + -->
<!--   xlab("") + -->
<!--   theme_pubr(border = T) + -->
<!--   guides(fill = F) -->
<!-- ``` -->


<!-- ### Explain expression differences betweeen high affinity special reporters -->
<!-- ```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE} -->
<!-- ## Rule of thumb for fitting models: e^n_features < n_observations -->

<!-- ## Subselect p53 only - let's only go for 2-2 combinations for now to simplify the model -->
<!-- cDNA_df_p53 <- cDNA_df %>% -->
<!--   filter(neg_ctrls == "No", positive_ctrl == "No", pDNA_library == "lib_2", tf == "P53", -->
<!--          spacing == 7, str_detect(affinity_id, "[10-12]{2}_"), position == 0, background == 1, n_sites == 4, -->
<!--          promoter == "mCMV", str_detect(condition, "MCF7_WT_Nutlin"), str_detect(condition, "PFT", negate = T)) %>% -->
<!--   dplyr::select(reporter_id, condition, reporter_activity_condition, -->
<!--                 affinity_pos1, affinity_pos2, affinity_pos3, affinity_pos4, motif_id, affinity_id) %>%  -->
<!--   mutate(reporter_activity = ave(reporter_activity_condition, motif_id, condition, FUN = function(x) mean(x)), -->
<!--          log_reporter_activity = log2(reporter_activity)) %>% -->
<!--   dplyr::select(-reporter_id, - reporter_activity_condition) %>% -->
<!--   unique() -->

<!-- cDNA_df_p53$low_aff_site <- apply(cDNA_df_p53 %>% dplyr::select(affinity_pos1, affinity_pos2, affinity_pos3, affinity_pos4), 1, max) -->
<!-- cDNA_df_p53$low_aff_site <- as.factor(cDNA_df_p53$low_aff_site) -->

<!-- ## Explain which affinities are essential at which position to boost activity -->
<!-- cDNA_df_p53 <- cDNA_df_p53[order(cDNA_df_p53$condition),] %>% -->
<!--   dplyr::select(-motif_id) -->
<!-- cDNA_df_p53$row <- rownames(cDNA_df_p53) -->

<!-- ## Fitting the model -->
<!-- x <- lm(reporter_activity ~ low_aff_site + affinity_id, cDNA_df_p53) -->

<!-- summ(x) -->
<!-- par(mfrow=c(2,2)) -->
<!-- plot(x, main = i) -->

<!-- # Get predicted activities per reporter -->
<!-- y <- data.frame(x$fitted.values, stringsAsFactors=FALSE) %>% rownames_to_column() -->
<!-- names(y) <- c("row", "reporter_activity_predicted") -->
<!-- y$id <- 1:nrow(y) -->
<!-- prediction <- data.frame("id" = 1:(nrow(cDNA_df_p53)), stringsAsFactors=FALSE) -->
<!-- prediction <- merge(prediction,y, all = T) %>%  -->
<!--   dplyr::select(-id) -->
<!-- cDNA_df_p53 <- merge(cDNA_df_p53, prediction) -->


<!-- # What are the individual weights? -->
<!-- weight <- data.frame(x$coefficients, stringsAsFactors=FALSE) %>% rownames_to_column() -->
<!-- names(weight) <- c("feature", "weight") -->
<!-- weight <- weight %>% filter(feature != "(Intercept)") -->
<!-- categorie <- c("affinity_pos", "n_sites", "highest_dist") -->
<!-- weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature) -->
<!-- weight$cond <- gsub(paste(weight$features, collapse = "|"), "", weight$feature) -->


<!-- # How many percent do the different tested features explain? -->
<!-- exp_variance <- anova(x) %>% rownames_to_column("feature") %>% -->
<!--   setnames("Sum Sq", "sum_sq") %>% -->
<!--   dplyr::select(feature, sum_sq) %>% -->
<!--   mutate(sum = sum(sum_sq)) %>% -->
<!--   mutate(rel_sum_sq = (sum_sq/sum)*100) %>% -->
<!--   setnames("rel_sum_sq", "percent") %>% -->
<!--   dplyr::select("feature", "percent") -->

<!-- exp_variance$group <- "no residual" -->
<!-- exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual" -->
<!-- exp_variance$percent[exp_variance$feature == "Residuals"] <- 100 - exp_variance$percent[exp_variance$feature == "Residuals"] -->
<!-- exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total")) -->


<!-- # Plotting -->
<!-- ggplot(exp_variance,  -->
<!--        aes(x = percent, y = reorder(feature, -percent), fill = group)) + -->
<!--   scale_fill_manual(values = c("#1B998B", "#2D3047")) + -->
<!--   geom_bar(stat = "identity") +  -->
<!--   xlab("Variance explained (%)") +  -->
<!--   ylab("") +  -->
<!--   labs(title = "Linear variance modelling",  -->
<!--        subtitle = "Expression variance predicted by features") + -->
<!--   theme_pubr(border = T) + -->
<!--   theme(text = element_text(size = 12)) + -->
<!--   guides(fill = F) -->

<!-- ggscatter(cDNA_df_p53, x = "reporter_activity", y = "reporter_activity_predicted", -->
<!--           add = "reg.line", -->
<!--           add.params = list(color = "#84A98C", fill = "lightgray"), -->
<!--           conf.int = TRUE, ylab = "Predicted reporter activity", size = 3, -->
<!--           xlab = "Average reporter activity") +  -->
<!--   theme_pubr(border = T)+ -->
<!--   stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 0, label.y = 3) -->

<!-- ggplot(weight,  -->
<!--        aes(x = reorder(feature, -weight), y = weight)) + -->
<!--   geom_bar(stat = "identity", width = 0.7, fill = colors_continous[2], color = "black") +  -->
<!--   ylab("weight - reporter activity") + -->
<!--   xlab("") + -->
<!--   theme_pubr(border = T) + -->
<!--   guides(fill = F) -->
<!-- ``` -->




<!-- ### Ridge/Lasso regression -->
<!-- ```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE} -->
<!-- # Train model -->
<!-- fit <- cva.glmnet(reporter_activity_condition ~ affinity_pos1*affinity_pos2*affinity_pos3*affinity_pos4, data=cDNA_df_p53) -->

<!-- # Get all parameters -->
<!-- get_model_params <- function(fit) { -->
<!--   alpha <- fit$alpha -->
<!--   lambdaMin <- sapply(fit$modlist, `[[`, "lambda.min") -->
<!--   lambdaSE <- sapply(fit$modlist, `[[`, "lambda.1se") -->
<!--   error <- sapply(fit$modlist, function(mod) {min(mod$cvm)}) -->
<!--   best <- which.min(error) -->
<!--   data.frame(alpha = alpha[best], lambdaMin = lambdaMin[best], -->
<!--              lambdaSE = lambdaSE[best], eror = error[best]) -->
<!-- } -->

<!-- x <- get_model_params(fit) -->

<!-- # Perform GLM -->
<!-- cDNA_dfMod <- glm(reporter_activity_condition ~ affinity_pos1*affinity_pos2*affinity_pos3*affinity_pos4, data=cDNA_df_p53, family = "gaussian") -->
<!-- print(cDNA_dfMod) -->

<!-- ``` -->





<!-- ### Random forest implementation -->
<!-- ```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE} -->
<!-- # Select features and activity to predict -->
<!-- cDNA_df_p53 <- cDNA_df %>% -->
<!--   filter(neg_ctrls == "No", positive_ctrl == "No", str_detect(affinity_id, "[1,3,5,7,9]_"), pDNA_library == "lib_2", -->
<!--          promoter == "mCMV", str_detect(condition, "MCF7_WT_Nutlin")) %>% -->
<!--   dplyr::select(reporter_id, condition, reporter_activity_condition,  -->
<!--                 affinity_pos1,spacing, background) %>% -->
<!--   mutate(spacing = as.factor(spacing), -->
<!--          background = as.factor(background), -->
<!--          reporter_activity_condition = log2(reporter_activity_condition)) %>% -->
<!--   unique() -->

<!-- cDNA_df_p53 <- cDNA_df %>% -->
<!--   filter(neg_ctrls == "No", positive_ctrl == "No", str_detect(reporter_id, "random", negate = T), pDNA_library == "lib_2", -->
<!--          str_detect(condition, "MCF7_WT"), str_detect(condition, "PFT", negate = T), tf == "P53") %>% -->
<!--   dplyr::select(reporter_id, spacing, affinity_pos4, background, promoter) %>% -->
<!--   mutate(spacing = as.factor(spacing), -->
<!--          affinity_pos4 = as.factor(affinity_pos4)) %>% -->
<!--   unique() -->

<!-- cDNA_df_p53 <- cDNA_df_p53[cDNA_df_p53$reporter_id %in% nutlin_sensitive$reporter_id,] -->

<!-- nutlin_sensitive <- nutlin_sensitive[nutlin_sensitive$reporter_id %in% cDNA_df_p53$reporter_id,] -->

<!-- cDNA_df_p53 <- merge(cDNA_df_p53, nutlin_sensitive)  -->


<!-- cDNA_df_p53 <- cDNA_df_p53 %>% -->
<!--   mutate(sensitive = as.factor(sensitive)) -->


<!-- # Set random seed to make results reproducible: -->
<!-- set.seed(4125) -->
<!-- # Calculate the size of each of the data sets: -->
<!-- data_set_size <- floor(nrow(cDNA_df_p53)/2) -->
<!-- # Generate a random sample of "data_set_size" indexes -->
<!-- indexes <- sample(1:nrow(cDNA_df_p53), size = data_set_size) -->
<!-- # Assign the data to the correct sets -->
<!-- training <- cDNA_df_p53[indexes,] -->
<!-- validation1 <- cDNA_df_p53[-indexes,] -->

<!-- # Perform random forest analysis -->
<!-- rf_features <- randomForest(sensitive ~ spacing + affinity_pos4 + background + promoter,  -->
<!--                   data = training, ntree=500, mtry = 8, importance=TRUE) -->

<!-- # Evaluate the model -->
<!-- predictions <- as.data.frame(predict(rf_features, validation1, type = "prob")) -->
<!-- predictions_df <- cbind(predictions, validation1) -->
<!-- predictions_df$sensitive_2 <- 0 -->
<!-- predictions_df$sensitive_2[predictions_df$sensitive == "Yes"] <- 1 -->

<!-- # validation1$prediction <- pred -->
<!-- # ggscatter(validation1, x = "sensitive", y = "prediction", -->
<!-- #           add = "reg.line", -->
<!-- #           add.params = list(color = "blue", fill = "lightgray"), -->
<!-- #           conf.int = TRUE, ylab = "Predicted log2-activity", size = 0.5, -->
<!-- #           xlab = "Average log2-activity") +  -->
<!-- #   theme_pubr() +  -->
<!-- #   stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 1.5, label.y = 3) -->

<!-- roc_rf_test <- roc(response = predictions_df$sensitive_2, predictor = predictions_df$Yes) -->
<!-- roc_rf_train <- roc(response = training$sensitive, predictor = rf_features$votes[,2]) -->

<!-- plot(roc_rf_test, col="#264653", lwd=3, main="ROC curve RF") -->
<!-- plot(roc_rf_train, col = "#2a9d8f", add = TRUE, label = T) -->
<!-- auc(roc_rf_test) -->


<!-- # Extract features and plot their importance -->
<!-- feature_df <- data.frame(importance(rf_features), stringsAsFactors=FALSE) %>% rownames_to_column("feature") -->


<!-- plot_ly(feature_df, x = ~feature, y = ~X.IncMSE, type = 'bar', -->
<!--              marker = list(color = '#D6D5C9', -->
<!--                            line = list(color = 'rgb(8,48,107)', width = 1.5))) %>%  -->
<!--   layout(title = "Random forest - factor importance plot (P53-mCMV reporters)", -->
<!--          xaxis = list(title = "Feature"), -->
<!--          yaxis = list(title = "%IncMSE")) -->

<!-- # Variable importance with DALEX -->
<!-- library(DALEX) -->
<!-- explained_rf <- explain(rf_features, data=validation1 %>% dplyr::select(-reporter_id, -sensitive), y=validation1$sensitive) -->

<!-- # Get the variable importances -->
<!-- varimps = feature_importance(explained_rf, type='raw') -->

<!-- ggplot(varimps %>% filter(variable != "_full_model_", variable != "_baseline_"), aes(y = dropout_loss, x = variable)) + -->
<!--     geom_beeswarm()+ -->
<!--     geom_boxplot(alpha = 0.4)+ -->
<!--   theme_pubr()+ -->
<!--     ylab("dropout loss (RMSE)") -->
<!-- ``` -->





<!-- ### Random forest implementation - explain affinity position importance -->
<!-- ```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE} -->
<!-- # Select features and activity to predict -->
<!-- cDNA_df_p53 <- cDNA_df %>% -->
<!--   filter(neg_ctrls == "No", positive_ctrl == "No",  -->
<!--          spacing == "7", background == "1", str_detect(affinity_id,"[7-9]"), pDNA_library == "lib_2", -->
<!--          promoter == "mCMV", str_detect(condition, "MCF7_WT_Nutlin")) %>% -->
<!--   dplyr::select(reporter_id, condition, reporter_activity_condition,  -->
<!--                 affinity_pos1, affinity_pos2, affinity_pos3, affinity_pos4) %>%  -->
<!--   mutate(reporter_activity_condition = log2(reporter_activity_condition)) %>% -->
<!--   unique() -->



<!-- # Set random seed to make results reproducible: -->
<!-- set.seed(17) -->
<!-- # Calculate the size of each of the data sets: -->
<!-- data_set_size <- floor(nrow(cDNA_df_p53)/2) -->
<!-- # Generate a random sample of "data_set_size" indexes -->
<!-- indexes <- sample(1:nrow(cDNA_df_p53), size = data_set_size) -->
<!-- # Assign the data to the correct sets -->
<!-- training <- cDNA_df_p53[indexes,] -->
<!-- validation1 <- cDNA_df_p53[-indexes,] -->

<!-- # Perform random forest analysis -->
<!-- rf_features <- randomForest(reporter_activity_condition ~ affinity_pos1 + affinity_pos2 + affinity_pos3 + affinity_pos4,  -->
<!--                   data = training, ntree=100, mtry = 2, importance=TRUE) -->

<!-- # Evaluate the model -->
<!-- pred = predict(rf_features, newdata=validation1[,c(-1,-2,-3)]) -->
<!-- validation1$prediction <- pred -->
<!-- ggscatter(validation1, x = "reporter_activity_condition", y = "prediction", -->
<!--           add = "reg.line", -->
<!--           add.params = list(color = "blue", fill = "lightgray"), -->
<!--           conf.int = TRUE, ylab = "Predicted log2-activity", size = 0.5, -->
<!--           xlab = "Average log2-activity") +  -->
<!--   theme_bw() +  -->
<!--   stat_cor(aes(label = ..r.label..), method = "pearson", label.x = 1.5, label.y = 3) -->

<!-- # Extract features and plot their importance -->
<!-- feature_df <- data.frame(importance(rf_features), stringsAsFactors=FALSE) %>% rownames_to_column("feature") %>% -->
<!--   dplyr::select(-IncNodePurity) -->


<!-- plot_ly(feature_df, x = ~feature, y = ~X.IncMSE, type = 'bar', -->
<!--              marker = list(color = '#D6D5C9', -->
<!--                            line = list(color = 'rgb(8,48,107)', width = 1.5))) %>%  -->
<!--   layout(title = "Random forest - factor importance plot (P53-mCMV reporters)", -->
<!--          xaxis = list(title = "Feature"), -->
<!--          yaxis = list(title = "%IncMSE")) -->

<!-- # Variable importance with DALEX -->
<!-- library(DALEX) -->
<!-- explained_rf <- explain(rf_features, data=training[,c(-1,-2,-3)], y=training$reporter_activity_condition) -->

<!-- # Get the variable importances -->
<!-- varimps = variable_importance(explained_rf, type='raw') -->

<!-- ggplot(varimps %>% filter(variable != c("_full_model_","_baseline_")), aes(y = variable, x = dropout_loss)) + -->
<!--   geom_quasirandom()+ -->
<!--   geom_boxplot(alpha = 0.4)+ -->
<!--   xlim(0.2,0.8)+ -->
<!--   xlab("dropout loss (RMSE)")+ -->
<!--   theme_bw() -->
<!-- ``` -->



# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

